A conditional permutation-based approach to test confounder and center effects in machine learning models

### 2021-10-25 manuscript.tex

Simulations with non-normal data showed that the proposed methods are remarkably robust to the violation of normality assumptions for the target variable $\y$ and the confounder $\c$ and, as expected, completely non-parametric regarding the distribution of the predictions $\yhat$). The six simulation cases of normality violation, shown on Figure \ref{fig:sim-non-normal}, involved various \emph{sinh-arcsinh} distributions, form normal (Fisher's kurtosis: 0, skewness) to severely skewed and heavy tailed (kurtosis: 42, skewness: 6).
The results show that the violation of normality - as expected - increases the type I error rate, but the increase - depending on the application - might not be severely problematic up to kurtosis and skewness values as high as 5 and 2 respectively. Furthermore, in terms of type I error rate, normality violation becomes less problematic in case of high predictive performance (high association between the predicted and observed values).

>>>>

Simulations with non-normal data showed that the proposed methods are remarkably robust to the violation of normality assumptions for the target variable $\y$ and the confounder $\c$ and, as expected, completely non-parametric regarding the distribution of the predictions $\yhat$). The six simulation cases of normality violation, shown on Figure \ref{fig:sim-non-normal}, involved various \emph{sinh-arcsinh} distributions, from normal (Fisher's kurtosis: 0, skewness: 0) to severely skewed and heavy tailed (kurtosis: 42, skewness: -6).
The results show that the violation of normality - as expected - increases the type I error rate, but the increase - depending on the application - might not be severely problematic up to kurtosis and skewness values as high as 5 and 2 respectively. Furthermore, in terms of type I error rate, normality violation becomes less problematic in case of high predictive performance (high association between the predicted and observed values).
